---
title: "Lab03"
author: "Haoran Zhang"
date: "2021/9/10"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Read in the data
First download and then read in with data.table:fread()

```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("met_all.gz")
```

## 2. Check the dimensions, headers, footers. How many columns, rows are there?
```{r}
dim(met)
```
```{r}
head(met)
```

```{r}
tail(met)
```

## 3. Take a look at the variables.

```{r}
str(met)
```
## 4. Take a closer look at the key variables.

```{r}
table(met$year)
```

```{r}
table(met$day)
```

```{r}
table(met$hour)
```

```{r}
summary(met$temp)
```

```{r}
summary(met$elev)
```
```{r}
summary(met$wind.sp)
```
Replace elevations with 9999 as NA.
```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```
We also have the issue of the minimum temperature being -40C, so we should remove those observations.
```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

## 5. Check the data against an external data source.

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

## 6. Calculate summary statistics

```{r}
elev <- met[elev==max(elev)]
summary(elev)
```

```{r}
cor(elev$temp, elev$wind.sp, use="complete")
```

```{r}
cor(elev$temp, elev$hour, use="complete")
```

```{r}
cor(elev$wind.sp, elev$day, use="complete")
```

```{r}
cor(elev$wind.sp, elev$hour, use="complete")
```

```{r}
cor(elev$temp, elev$day, use="complete")
```

## 7. Exploratory graphs

```{r}
hist(met$elev, breaks=100)
```

```{r}
hist(met$temp)
```

```{r}
hist(met$wind.sp)
```

Look at where the weather station with highest elevation is located.
```{r}
library(leaflet)
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

Look at the time series of temperature and wind speed at this location. For this we will need to create a date-time variable for the x-axis.
```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

With the date-time variable we can plot the time series of temperature and wind speed.
```{r}
plot(elev$date, elev$temp, type='l')
```

```{r}
plot(elev$date, elev$wind.sp, type='l')
```
